Interactive Filter and Display of Hillary Clinton's Emails: \\A Cautionary Tale of Metadata

### 2017-03-31 VAST_submission/vgtc_journal_latex/paper.tex
Added zero dates
Added my edits

The greater the number of emails between Clinton and the correspondent, the wider is the spoke.  Whenever the difference in volume is great enough the correspondents will also be separated into two groups.  Those with the greatest correspondence will have shorter spokes and hence be visually ``closer'' to Clinton.   As the display shows, Clinton's closest inner circle are her closest aides Huma Abedin, Cheryl Mills, and Jake Sullivan. \TODO{Chris, can you be more precise about the algorithm?}
\subsection{Email volume}
\label{sect:Displays:volume}
Figure~\ref{fig:VolumeAll} has two lines, the first is a line showing the total volume of emails for each day, and the second shows the number of emails satisfying the classification filter criteria for a given day. This allows users to view which days have an exceptionally high or low number of classified conversations. Returning to the motivation of hypothesis generation, referencing world events on any days of interest provides a number of interesting insights on how public figures like Clinton react and respond to world events.

@@

\subsection{Term frequency and TFIDF}
\label{sect:Displays:TFIDF}
The 20 terms (stemmed words excluding stopwords) which appear in the greatest number of the emails selected are displayed as those having ``Highest Frequency''.  
\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{HighFreqAll} &
$~~~~$ &
\includegraphics[width=0.4\linewidth]{HighTFIDFAll} 
\end{tabular}
\end{center}
\end{figure}
Those terms which appear frequently within some emails but less frequently across emails will have a high tf-idf (term frequency - inverse document frequency) score.  The top 20 scoring of these in the selected emails are shown in the ``TFIDF'' display. \TODO{Chris, check that I have described these two measures correctly}
The top 20 shown here are for all emails in the corpus; these will change depending on the filtering.

@@

\subsection{Highest peak}
Within the plot of email volume over time, one very large peak stands out.
Historically important,  possibly had an impact on the 2016 US presidential election.

>>>>

The greater the number of emails between Clinton and the correspondent, the wider is the spoke. Spoke width is scaled linearly relative to the greatest number of emails for all correspondents in the selection, with a constant minimum width for visibility.  The correspondents are also separated into two groups by the greatest difference in ordered email volumes.  Those which appear before this difference when all are ordered by decreasing email volume have nodes placed at half the distance from the centre. will have shorter spokes and hence be visually ``closer'' to Clinton.   As the display shows, Clinton's closest inner circle are her closest aides Huma Abedin, Cheryl Mills, and Jake Sullivan. \TODO{Chris, can you be more precise about the algorithm?}
\subsection{Email volume}
\label{sect:Displays:volume}
Figure~\ref{fig:VolumeAll} has two lines, the first is a line showing the total volume of emails for each day, and the second shows the number of emails satisfying the classification filter criteria for a given day. This allows users to view which days have an exceptionally high or low number of classified conversations. Returning to the motivation of hypothesis generation, referencing world events on any days of interest provides a number of interesting insights on how public figures like Clinton react and respond to world events.

@@

\subsection{Term frequency and TFIDF}
\label{sect:Displays:TFIDF}
The 20 terms (stemmed words excluding stopwords) which appear in the greatest number of the emails selected are displayed as those having ``Highest Frequency''.  
\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.4\linewidth]{HighFreqAll} &
$~~~~$ &
\includegraphics[width=0.4\linewidth]{HighTFIDFAll} 
\end{tabular}
\end{center}
\end{figure}
Those terms which appear frequently within some emails but less frequently across emails will have a high tf-idf (term frequency - inverse document frequency) score.  The top 20 scoring of these in the selected emails are shown in the ``TFIDF'' display.
The top 20 shown here are for all emails in the corpus; these will change depending on the filtering.

@@

\subsection{Highest peak}
Within the plot of email volume over time, one very large peak stands out in the later half of 2011. Using the time filtering function the date of this peak can be identified as August 21, 2011. A quick internet search of this date reveals that it was the start of the Battle of Tripoli, the final battle in the 2011 Libyan Civil War in which the NATO-backed National Transitional Council \cite{NATONTCSupport,LibyaContactGroup} toppled the regime of Muammar Gaddafi. \cite{BattleforTripoli,AlJazeeraTripoli} This suggests that the spike in emails was a response by the State Department as a whole to the news of the National Transitional Council's final attacks on Tripoli. The tf-idf and term frequencies over this time frame support the suggestion that this event was the focus of most of the communication.

Terms included in this selection such as "tnc" (the abbreviation for the National Transitional Council rebels used by Clinton), "jalil" (the head of the National Transitional Council \cite{Jalil}), "libya", "tripoli", and "qadhafi" clearly refer to the Libyan Civil War. Of greater interest are some of the less clearly connected terms, such as "davutoglu," "tarhouni," and "jupp." Searching for these terms gives some immediate and obvious connections. "Davutoglu" appears to refer to Ahmet Davutoglu, the Turkish Minister of Foreign Affairs at the time. "Jupp" seems to be a reference to Alain Juppe, the French Minister of Foreign Affairs. Finally, "tarhouni" likely refers to Ali Tarhouni, the Minister for Oil and Finance within the National Transitional Council. The mention of these particular names gives insight into the individuals the State Department likely viewed as most significant during the Battle of Tripoli and transition of the Libyan government.

\subsection{No email sent by Clinton}

Using 1 month window slider and looking for flat spots at zero

Dates:
\begin{itemize}
\item None until mid April 2009
\item June 8 or 9, June 15 to about the 20th, 2009
\item July 17-20,  around 27, 2009 
\item Day or 2 around Oct 13, 14, 2009
\item April 10-15 or so? 2011
\item one or two days late August, 2011. (no mail)
\item end of march first half of April 2012
\item August 27-30  2012 
\item low first week of September, 2012
\item 2 weeks end of october, up to Nov 10 approx 2012
\item Dec 7- 17? 2012,  Again about Dec 25 2012 to Jan 1, 2013
\end{itemize}
